Analysis-ready VCF at Biobank scale using Zarr

### 2025-01-08 paper.tex
Add BAI/VCF POS paragraphs
- exemplify old indexing formats that are ill suited for large genomes
- motivate why the VCFs were partitioned
- mention VCF POS limitation and how this could affect other genomes
  with large chromosomes

%% FIXME: reference regarding VCF POS?
The spruce data set poses an interesting challenge to the VCF storage
format due to the large chromosome sizes (>1Gbp). The VCF POS field is
encoded in 32-bit format (corresponding to a maximum of approximately
537Mbp), which mandates splitting up larger chromosomes into chunks,
adding additional complexity to downstream analyses. Zarr does not
impose this limitation, and we therefore demonstrate the versatility
of Zarr by updating contig information and the coordinate system for
the spruce dataset.

>>>>

Large genomes from non-model organisms pose particular problems for
analysis workflows that often are tailored for human conditions. For
instance, the binary Sequence Alignment/Map index format BAI has a 512
Mbp limit, but spruce chromosomes range in size from 1.0 to 1.7 Gbp.
The spruce BAM files used for variant calling were BAI indexed, which
motivated the partitioning into multiple VCF files in the first place,
adding complexity to downstream processing.

A similar limitation is imposed by the VCF POS field which is encoded
in 32-bit format and can hold chromomosomes up to 2.14 Gbp. Although
spruce does not have chromosomes that exceed the limit, other species
do, such as the recently sequence mistletoe, with chromosomes larger
than 8 Gbp. Zarr does not impose this limitation, and we therefore
demonstrate the versatility of Zarr by updating contig information and
the coordinate system for the spruce dataset. This updating procedure
could be applied to any other species.
